\begin{thebibliography}{99}

\bibitem{Aloni} Aloni, M. and van Rooij, R. (2002), ``The Dynamics
of Questions and Focus", SALT 2002.

\bibitem{Dretske} Dretske, F. (1970), ``Epistemic Operators", \emph{The Journal of Philosophy},
Vol. 67, No. 24 (Dec. 24, 1970), 1007-1023.

\bibitem{GS} Groenendijk, J. and Stokhof M. (1984), \emph{Studies
in the Semantics of Questions and the Pragmatics of Answers}. PhD
dissertation, University of Amsterdam.

\bibitem{Hamblin} Hamblin, C.L. (1973), ``Questions in Montague
English", \emph{Foundations of Language}, 10, 41-53.

\bibitem{Hasp} Haspelmath, M. (2000), ????

\bibitem{Hawthorne} Hawthorne, J. (2004), \emph{Knowledge and
Lotteries}, Oxford University Press.

\bibitem{Heim} Heim, I. (1994), ``Interrogative Semantics and Karttunen's
Semantics for \emph{Know}", in R. Buchalla and A. Mittwoch (eds)
\emph{Proceedings of IATL 1}, 128-144. Akademon, Jerusalem.

\bibitem{K} Karttunen, L. (1977), ``Syntax and Semantics of
Questions", \emph{Linguistics and Philosophy}, 1, 3-44.

\bibitem{chan} Romero, M. and Han C-H. (2003), ``Focus, Ellipsis and
the Semantics of Alternative Questions", in \emph{Empirical Issues
in Formal Syntax and Semantics 4}. C. Beyssade, O. Bonami, P.
Cabredo Hofherr, F. Corblin (eds), Presses Universitaires de
Paris-Sorbonne, Paris, 291-307.

\bibitem{Schaffer} Schaffer, J. (2005), ``Knowing the Answer",
Penultimate Draft, forthcoming in \emph{Philosophy and
Phenomenological Research}.


\end{thebibliography}
